---
title: "Covid91 vaccine study Final2024"
author: "Giselle Gonzalez!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

 $H_0$: The covid921 vaccine is not effective in males  
 $H_A$: The covid921 vaccine is effective in males 



###  Methods   
 Treatment is a catergorical variable 
 Sex is catagorical variable 
 I use cat~CAT helper file 


###  Descriptive results


### Graphical results

```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected,data=males)
barchartGC(~infected + treatment,data=males,type="percent")

```

 This shows that males who got the vaccine are less likely to get covid than people who got the placebo.
 
### Nummerical results 
```{r}
table1 <- xtabs(~treatment + infected,data=males)
rowPerc(table1)
colPerc(table1)
```
 This shows that 74% of males who took the placebo got covid19, and 26 percent males who got the vaccine  got covid19
 
### Inferential results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
 Both the chi and fisher have The same p-value which  is 0.000000000000002 which means its lower than 0.05   
The odds ratio is 28.1:1  means that the placebo is 2.81 which  people are more likely to get covid 
out of the 1 who took the vaccine 

### Conclusion 

## Females

$H_0$: The covid921 vaccine is not effective in females 
 $H_A$: The covid921 vaccine is effective in females 

###  Methods   
 Treatment is a catergorical variable 
 Sex is catagorical variable 
 I use cat~CAT helper file 


###  Descriptive results


### Graphical results

```{r}
females=filter(FinalData,FinalData$sex=="F")
barchartGC(~treatment + infected,data=females)
barchartGC(~infected + treatment,data=females,type="percent")

```

 This shows that males who got the vaccine are less likely to get covid than people who got the placebo.
### Nummerical results 
```{r}
table1 <- xtabs(~treatment + infected,data=females)
rowPerc(table1)
colPerc(table1)
```
 This shows that 74% of females who took the placebo got covid19, and 26 percent females who got the vaccine  got covid19
 
### Inferential results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Both the chi and fisher have The same p-value which  is 0. which means its lower than 0.05 the odds ratio is 1.85:1 the placebo is 1.85 and the vaccine is 1 wich it is more likey 

## LGBTQ

 $H_0$: The covid921 vaccine is not effective in lgbtq  
 $H_A$: The covid921 vaccine is effective in lgbtq



###  Methods 

Treatment is a catergorical variable 
 Sex is catagorical variable 
 I use cat~CAT helper file 

###  Descriptive results


### Graphical results



## Druggies


# Overall Results and Conclusions