---
title: "Covid91 vaccine study Final2024"
author: "Giselle Gonzalez!"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="finalSumr2024.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

## Males

 $H_0$: The covid921 vaccine is not effective in males  
 $H_A$: The covid921 vaccine is effective in males 



###  Methods   
 Treatment is a catergorical variable 
 the sex(males) is a catagorical variable 
 I use cat~cat helper file 


###  Descriptive results


### Graphical results

```{r}
males=filter(FinalData,FinalData$sex=="M")
barchartGC(~treatment + infected,data=males)
barchartGC(~infected + treatment,data=males,type="percent")

```

According to the graph, men who got the vaccine are at a reduced risk of getting COVID compared to individuals who received the placebo
 
### Nummerical results 
```{r}
table1 <- xtabs(~treatment + infected,data=males)
rowPerc(table1)
colPerc(table1)
```

 The data shows that among males, 74% who took the placebo contracted COVID-19, compared to only 26% of those who received the vaccine.
 
### Inferential results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
 
 The chi-square test p-value is 0.0000000000000002  , and the Fisher's exact test p-value is 
0.000000000000002, which is significantly less than 0.05. The odds ratio is 2.82:1, meaning that people who took the placebo are 2.82 times more likely to contract COVID-19 compared to those who received the vaccine.

### Conclusion 

According to the graph, men who got vaccinated are at a significantly lower risk of contracting COVID-19 compared to those who received the placebo.   
The data shows that 74% of males who took the placebo contracted the virus, while only 26% of those who got the vaccine did. The chi-square test p-value is 0.0000000000000002, and the Fisher's exact test p-value is 0.000000000000002 which is much less than 0.05, indicating strong statistical significance so we reject the null.   
The odds ratio of 2.82:1 means that those who received the placebo are 2.82 times more likely to develop COVID-19 compared to those who were vaccinated.

## Females

$H_0$: The covid921 vaccine is not effective in females   
 $H_A$: The covid921 vaccine is effective in females 

###  Methods   
 Treatment is a catergorical variable   
 sex (females)is catagorical variable   
 I use cat~CAT helper file 


###  Descriptive results


### Graphical results

```{r}
females=filter(FinalData,FinalData$sex=="F")
barchartGC(~treatment + infected,data=females)
barchartGC(~infected + treatment,data=females,type="percent")

```

This indicates that females who were vaccinated are less likely to contract COVID-19 than those who were given the placebo.
 .
### Nummerical results 
```{r}
table1 <- xtabs(~treatment + infected,data=females)
rowPerc(table1)
colPerc(table1)
```
 This shows that 65% of females who took the placebo got covid19, and 35 percent females who got the vaccine  got covid19
 
### Inferential results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

 With a chi-square p-value of 0.0000002 and a Fisher’s exact test p-value of 0.0000001, both of which are less than 0.05, we reject the null hypothesis.   
The odds ratio of 1.85:1 shows that females who took the placebo are 1.85 less likely to contract COVID-19 compared to those who were vaccinated 

### Conclusion 
This indicates that vaccinated females are less likely to contract COVID-19 than those who received the placebo.  
The data reveals that 65% of females who took the placebo contracted COVID-19, whereas 35% of vaccinated females contracted the virus. With a chi-square p-value of 0.0000002 and a Fisher’s exact test p-value of 0.0000001—both less than 0.05—we reject the null hypothesis.  
The odds ratio of 1.85:1 indicates that females who took the placebo are 1.85 times less likely to contract COVID-19 compared to those who were vaccinated.

## LGBTQ

 $H_0$: The covid921 vaccine is not effective in lgbtq  
 $H_A$: The covid921 vaccine is effective in lgbtq



###  Methods 

Treatment is a catergorical variable 
 sex (lgbtq)is catagorical variable 
 I use cat~CAT helper file 

###  Descriptive results

### Graphical results
```{r}
LGBTQ=filter(FinalData, LGBTQ=="gay")

```


```{r}
barchartGC(~infected+ treatment,data=LGBTQ)
barchartGC(~infected + treatment,data=LGBTQ, type="percent")

```

According to the graph, LGBTQ people who were vaccinated have a decreased risk of getting COVID-19 compared to those who were given a placebo.

### Nummerical results 
```{r}
table1 <- xtabs(~treatment + infected,data=LGBTQ)
rowPerc(table1)
colPerc(table1)
```

 This shows that 80% of lgbtq who took the placebo got covid19, and 20 percent of lgbtq who got the vaccine  got covid19 
### Inferential results
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The chi-squared pvalue is 0.00002 and the fisher pvalue is 0.0000007 which is lower than 0.05 meaning we reject the null. The odds ratio is 4.27:1, and the lgbtq who took the placebo is 4.27 less likely to get covid then the 1 who took the vaccine 

### Conclusion 
According to the graph, LGBTQ individuals who received the vaccine showed a reduced risk of contracting COVID-19 compared to those who received a placebo. Specifically, 80% of LGBTQ individuals who took the placebo contracted COVID-19, while 20% of those who received the vaccine contracted the virus. The chi-squared p-value is 0.00002 and the Fisher's p-value is 0.0000007, both of which are less than 0.05, indicating rejection of the null hypothesis. The odds ratio is 4.27:1, suggesting that LGBTQ individuals who took the placebo are 4.27 times less likely to contract COVID-19 compared to those who received the vaccine.

## Druggies

 $H_0$: The covid921 vaccine is not effective in drug users  
 $H_A$: The covid921 vaccine is effective in drug users

###  Methods   
 Treatment is a catergorical variable   
 sex-druguser is catagorical variable  
 I use cat~CAT helper file   

## Descriptive Results

### Graphical Results```{r}

```{r}
DruggiesData=filter(FinalData, DrugUser=="yes")
```


```{r}
barchartGC(~infected+treatment,data=DruggiesData)
barchartGC(~infected+treatment,data=DruggiesData, type="percent")
```

According to the graph, individuals who used the placebo among drug users were less likely to contract COVID-19 compared to those who received the vaccine


### Numerical Descriptive Results

```{r}
table4 <- xtabs(~treatment+ infected, data=DruggiesData)
rowPerc(table4)
colPerc(table4)
```

This data shows that 8.33 of drug users who took the placebo got covid19, and 91.67  of drugusers who got the vaccine  got covid19 

## Infrenital Results

```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```

The chi-squared pvalue is 0.0000000001 and the fisher pvalue is 0.000000000003 which is lower than 0.05 meaning we reject the null. The odds ratio is 0.087:1, and the drugusers who took the placebo is 0.08 less likely to get covid then the 1 who took the vaccine 


### Conclusion 
According to the graph, drug users who received the placebo were less likely to contract COVID-19 compared to those who received the vaccine. Specifically, 8.33% of drug users who took the placebo contracted COVID-19, while 91.67% of drug users who received the vaccine contracted the virus. The chi-squared p-value is 0.0000000001 and the Fisher's p-value is 0.000000000003, both of which are less than 0.05, indicating rejection of the null hypothesis. The odds ratio is 0.087:1, suggesting that drug users who took the placebo are 0.087 times less likely to contract COVID-19 compared to those who received the vaccine.





